---
title: "Referee Report"
output:
  pdf_document: 
    fig_height: 4
    fig_width: 7
    toc: yes
  html_notebook: 
    toc: no
    code_folding: none
  html_document: default
echo: FALSE
---

<style type="text/css">
.table {
    width: inherit;
    max-width: 100%;
    margin-bottom: 20px;
}
.math.display {
  font-size: 28px;
}
</style>


```{r global-options, include=FALSE}
library(tidyverse)
library(knitr) %>% suppressWarnings()
theme_set(theme_light())
source('referee_data.R')

name <- 'Kayser, Bill'
username <- NULL
password <- NULL

friendly_name <- str_split_fixed(name, ", ", 2) %>% rev() %>% str_c(collapse=" ")

db <- collect_ref_data(name, username, password)
all_games <- db$games %>% split(.$upcoming)

games <- all_games[["FALSE"]]
upcoming <- all_games[["TRUE"]]
options(knitr.kable.NA = '')
knitr::opts_chunk$set(fig.width=12, 
                      fig.height=8, 
                      fig.path='Figs/',
                      echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE)


```

## Summary

`r nrow(games)` game were recorded with **`r friendly_name`** on the crew.

```{r echo=F}
games.total <- nrow(games)
games.center <- nrow(filter(games, position == 'center'))
games.ar1 <- nrow(filter(games, position == 'ar1'))
games.ar2 <- nrow(filter(games, position == 'ar2'))
games.ar <- games.ar1 + games.ar2
games.solo <- nrow(filter(games, system == 'Single'))

```

* `r games.center` were Center assignments
* `r games.solo` were Solo Center assignments
* `r games.ar` were Assistant Referee assignments
* `r games.ar1` were AR1 assignments
* `r games.ar2` were AR2 assignments

```{r, fig.width=10, fig.height=4}
games_by_month <- games %>% 
  mutate(m = lubridate::month(time),
         y = lubridate::year(time),
         period = (y-2017)*12 + m,
         month = str_c(m,"/",y%%100),
         date = lubridate::my(month)) %>%
  select(period, month, date, org) %>%
#  group_by(period, org, month) %>%
#  summarize(count=n()) %>%
  as_tibble()

period.val <- min(games_by_month$period):max(games_by_month$period)
period.month <- str_c((period.val %% 12) + 1, '/', 17 + (period.val %/% 12))

games_by_month$month <- factor(games_by_month$month, period.month)

ggplot(games_by_month) +
  aes(x=date, fill=org) +
  geom_bar() +
  scale_x_date(date_breaks = '3 months', date_labels = '%m/%y') +
  xlab('Month') +
  ylab('Number of Assignments') +
  ggtitle('Assignment History')

```

## Center Games{#centers}

```{r echo=F}
by_level <- games %>% 
    filter(position == 'center') %>%
    group_by(gender, level) %>%
    summarize(games = n()) %>%
    as_tibble()
```

<style>
 div#boys {
    display: inline-block;
    margin: 4px 4px 4px 12px;
 }
 div#girls {
    float: left;
    margin: 4px;
 }
 div#centers {
    float: left;
    margin: 4px;
 }
 div#organizations {
    display: inline-block;
    margin: 4px 4px 4px 32px;
 }
</style>


```{r}
girls <- filter(by_level, gender == 'Girls') %>% select(-gender) 
boys <- filter(by_level, gender == 'Boys') %>% select(-gender) 
```

#### Girls (`r sum(girls$games)`){#girls}

```{r}
kable(girls)
```

#### Boys (`r sum(boys$games)`){#boys}

```{r}
kable(boys)
```

## Organizations

```{r}
table(games$org) %>% sort(TRUE) %>% kable(col.names=c('Org', 'No. of Games'))
```

## Teammates

These are all referees appearing in game reports as part of your crew.


```{r}

teammates <- games %>%
  select(-position) %>%
  pivot_longer(c('center', 'ar1', 'ar2', 'assessor'), 
               values_to = 'referee', 
               names_to = 'position') %>%
  filter(!is.na(referee) & referee != friendly_name & referee != "") %>%
  # pivot_wider(names_from = position,
  #             values_from = position,
  #             values_fill = NA) %>%
  arrange(date) %>%
  group_by(referee) %>%
  summarize(games = n(), 
            last_game = last(date),
            location = last(field),
            last_teams = last(str_c(home, ' vs ', away)),
            center = sum(position=='center'),
            ar1=sum(position=='ar1'),
            ar2=sum(position=='ar2'),
            assessor=sum(position=='assessor')) %>%
  mutate(CanonicalName = str_to_upper(referee))  %>%
  left_join(db$referees, by=c("CanonicalName")) %>%
  mutate(Age = as.numeric(floor( as.numeric(Sys.Date() - Birthdate, unit="days")/365.25))) %>%
  select(Full_Name=referee,
         Games=games,
         Last_Game=last_game,
         Location=location,
         Last_Teams=last_teams,
         Center=center,
         AR1=ar1,
         AR2=ar2,
         Assessor=assessor,
         RefereeType,
         Age)

```

### Top Five Partners

```{r}

teammates %>% 
    arrange(desc(Games), desc(Center)) %>%
    head(5) %>% kable()
```

### Complete Index by Name

```{r}
kable(teammates)
```