---
title: "NCSC Referee Report"
output:
  html_notebook: 
    toc: no
    code_folding: none
  pdf_document: 
    fig_height: 4
    fig_width: 7
    toc: yes
  html_document: default
---
```{r setup, include=F}
library(kableExtra)
library(scales)
library(ggplot2)
library(lubridate)
library(knitr) %>% suppressWarnings()
knitr::opts_chunk$set(echo=F)
theme_set(theme_light())

```

<style type="text/css">
.table {
    width: 100%;
    max-width: 100%;
    margin-bottom: 20px;
}
.math.display {
  font-size: 28px;
}
</style>

This report was created using data scraped of [Oregon Soccer Central](https://www.oregonsoccercentral.com) as well
as the registered referees spreadsheets that are distributed by the [Oregon Referee Committe](https://www.oregonreferee.com).

## Summary Information

```{r init, include=F}
source('lib/ncsc_refs.R')
session <- create_session('Kayser, Bill')

refs <- read_games(session) |>
    pivot_longer(all_of(c('Center', 'AR1', 'AR2')),
                 names_to="Position",
                 values_to="UID") |>
    mutate(Year=year(Date),
           Years=year(Date)) |>
#    filter(!is.na(UID)) |>
    pivot_wider(names_from=Years, values_from=Years, values_fn=length, names_prefix='GamesIn', values_fill=0) |>
    group_by(UID) |>
    summarize(
        Center.7v7 = sum(Game == '7v7'),
        Center.9v9 = sum(Game == '9v9' & Position == 'Center'),
        Center.11v11 = sum(Game == '11v11' & Position == 'Center'),
        Center = sum(Position == 'Center'),
        AR = sum(Position == "AR1" | Position == "AR2"),
        Fall = sum(Season == 'Fall'),
        Spring = sum(Season == 'Spring'),
        across(starts_with("GamesIn"), sum),
        Rank1 = sum(Rank == 1),
        Rank2 = sum(Rank == 2),
        Rank3 = sum(Rank == 3),
        RankOther = sum(Rank > 3),
        FirstYear = min(Year),
        LastYear = max(Year),
        Total=n(),
    )  |>
    left_join(x=osc_refs(session), by='UID') |>  # read_refs returns refs currently in the org
#    filter(ID != "") |>
    mutate(across(!(1:23), \(v) ifelse(is.na(v), 0, v))) |> # Write zeros into all the NAs introduced in the join.
    mutate(Current = Registered.2022,
           Recent = GamesIn2022 > 0,
           New = InitialYear >= 2022 & Total == 0,
           Inexperienced = Age < 18 & (Center.9v9 + Center.11v11) == 0)
```


```{r summary}
n.registered <- nrow(refs)
n.current <- sum(refs$Current)
n.current.active2022 <- sum(refs$Current & refs$GamesIn2022 > 0)
n.active2022 <- sum(refs$GamesIn2022 > 0)
n.active.since.2021 <- sum(refs$GamesIn2022 > 0 | refs$GamesIn2021 > 0)
```

Right now there are **`r n.registered`** referees listed in our organization. 

Of those, **`r n.current`** have 2022 registrations completed.  Of currently registered refs, **`r n.current.active2022`** have already worked for us in 2022.

In total, there have been **`r n.active.since.2021`** referees who were active in 2021 or 2022.

Of the `r sum(refs$Recent)` refs who have worked at least one game in 2022, `r sum(refs$Recent & refs$Gender == 'F')` identify as female.

Here is a breakdown of the crew officiating games since 2017.  Note there were no games in 2020 because of the pandemic.  Also note that this only includes referees who were registered in 2019 or later.

```{r gender}
margins <- filter(refs, LastYear >= 2017) |>
    select(Gender, Age, starts_with('GamesIn')) |>
    pivot_longer(cols=starts_with('GamesIn'), values_to="Games") |>
    mutate(Year = str_match(name, "GamesIn(....)")[,2] |> as.integer()) |>
    select(-name)

group_by(margins, Gender, Year) |>
    summarize(Games = sum(Games)) |>
    ggplot() +
        aes(fill=Gender, x=as.factor(Year), y=Games) +
    geom_col() +
    ggtitle("Officiating Positions by Gender",
            "Count of positions (Center, ARs) paid by NCSC") +
    xlab(NULL) + ylab(NULL)

```

## Data Doctor

I ran these scripts to update data on [OregonSoccerCentral.com](http://www.oregonsoccercentral.com).  The changes have already been applied so there shouldn't be any more adjustments needed.

--------

For refs that have actually worked games for us, the `FirstYear` and `LastYear` fields will reflect games actually worked. 
If our manually entered `InitialYear` and `CurrentYear` don't align with actual games worked, we can fix that.

```{r fixit1, message=FALSE, warning=FALSE, include=FALSE}
# Find refs who have worked games where the initial year does not line up, or the current year needs to be moved up.
misaligned.refs <- 
    filter(refs,
           (Total > 0) & 
               (is.na(Org.InitialYear) | 
                    is.na(Org.CurrentYear) |
                    FirstYear != Org.InitialYear | 
                    LastYear > Org.CurrentYear)) |>
    mutate(Prev.InitialYear = Org.InitialYear,
           Prev.CurrentYear = Org.CurrentYear,
           Org.InitialYear = FirstYear,
           Org.CurrentYear = ifelse(is.na(Org.CurrentYear), LastYear, pmax(Org.CurrentYear, LastYear)))

updateRefDetails(session, misaligned.refs)
```

Refs needing adjustment: `r nrow(misaligned.refs)`

---------

Some refs have no Current or Initial Year for the org because when they were added these values were not set.

```{r noyears, include=FALSE}
never.worked <- filter(refs, 
                       is.na(Org.InitialYear) & is.na(Org.CurrentYear))

```
There are `r nrow(never.worked)` such refs registered with us currently.

----------

Some refs had their Current Year entered but not the initial year.  Maybe when setting up the ref initially, we just didn't bother setting this field.  For these situations, I'll assume the Initial Year is the same as the Current Year.

```{r fixit2, include=FALSE}
# Find refs who have worked games where the initial year does not line up, or the current year needs to be moved up.
refs.InitialYear.missing <- 
    filter(refs,
           is.na(Org.InitialYear) & !is.na(Org.CurrentYear)) |>
    mutate(Org.InitialYear = Org.CurrentYear)


updateRefDetails(session, refs.InitialYear.missing)
 
```

Refs needing adjustment: `r nrow(refs.InitialYear.missing)`

-------

Some refs did not get their ranks set.  These will show up as NA or 0.  Let's default the Center rank to 1 (7v7) and AR to 2 (9v9)

```{r fixit3, include=FALSE}
# Find refs who have worked games where the initial year does not line up, or the current year needs to be moved up.
refs.Ranks.missing <- 
    filter(refs,
           is.na(Rank.Center) | is.na(Rank.AR) | Rank.Center == 0 | Rank.AR == 0) |>
    mutate(Rank.Center = ifelse(is.na(Rank.Center) | Rank.Center == 0, 1, Rank.Center),
           Rank.AR = ifelse(is.na(Rank.AR) | Rank.AR == 0, 1, Rank.AR))

updateRefDetails(session, refs.Ranks.missing)
 
```

Refs needing adjustment: `r nrow(refs.Ranks.missing)`


## Refs by Experience

Here is the list of the top 25 refs by number of games.

```{r toprefs}

arrange(refs, desc(Total)) |>
    head(25) |>
    select(Name = FullName, Age, `Total Games`=Total, `Centers`=Center, 
           `Centers 11v11` = Center.11v11, `Reg 22?`=Registered.2022, 
           `Most Recent Year`=LastYear, `Last Logged In`=LastLogin) |>
    kbl() |>
    # column_spec(7, width = "5cm") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = T) 

```

## Spring '22 Summary

These are all the refs that did at least one game in Spring of '22.

```{r spring22}
spring22 <- season_summary(session, begin_date=mdy("3/1/2022"), end_date=mdy("6/14/2022")) |>
    arrange(desc(Total)) |>
    filter(Total > 0) |>
    mutate(Rank = str_c(Rank.Center, '/', Rank.AR)) 
select(spring22,
       Total,
       Center,
       AR,
       FullName,
       Age,
       Gender,
       City,
       Rank) |>
    kbl() |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = T) 


```

```{r seasonstats}


```
## Distribution Lists

### Refs That Might Be Interested in Training

```{r training}

refs.training <- 
    filter(refs,
           Grade == 8 & 
               (Age <= 16 |
                    InitialYear >= 2021)
    ) |>
    arrange(Age, LastName) 


select(refs.training,
       FullName,
       Age,
       InitialYear,
       CurrentYear,
       Org.InitialYear,
       Total) |>
    kbl() |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = T) 

```

```
`r generateEmailTo(refs.training)`
```

# Need Refs for Upcoming Games

### Need Centers for This Weekend

These are refs who are registered, have Centered 9v9 or above, and have done at least 1 game in 2021

```{r centersneeded}
immediate.centers <- filter(refs, 
                            Registered.2022 & LastYear >= 2021 & Center > 0) |>
    arrange(desc(Center.11v11), desc(Center.9v9), desc(Center.7v7)) 

select(immediate.centers,
       Name=FullName, 
       Age, 
       `Rank`=Rank.Center, 
       `11v11 Center`=Center.11v11, 
       `9v9 Center`=Center.9v9, 
       `7v7 Center`=Center.7v7,
       `Spring Games`=Spring, 
       Cell) |> 
    kbl() |>
    #    column_spec(7, width = "5cm") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = T) 

```

```
`r generateEmailTo(immediate.centers)`
```

### Need ARs for Upcoming Games

```{r need-ars}
immediate.ars <- filter(refs, 
                        Registered.2022 & LastYear >= 2021 &  !(UID %in% immediate.centers$UID)) |>
    arrange(desc(AR)) 

select(immediate.ars,
       Name=FullName, 
       Age, 
       `AR Rank`= Rank.AR, 
       `AR Assignments`= AR,
       `Spring Games`=Spring, 
       Cell) |>
    kbl() |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = T) 
```

```
`r generateEmailTo(immediate.ars)`
```
