---
title: "Oregon Referee Report"
always_allow_html: true
output:
  html_notebook: 
    code_folding: none
    theme: readable
    css: referee_style.css
    fig_width: 12
    fig_height: 6
    fig_caption: yes
  html_document: 
    code_folding: none
    theme: readable
    css: referee_style.css
  pdf_document: 
    fig_height: 8
    fig_width: 12
echo: FALSE
---

<style type="text/css">
.table {
    width: inherit;
    max-width: 100%;
    margin-bottom: 20px;
}
.math.display {
  font-size: 28px;
}
</style>


```{r global-options, include=FALSE}
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(scales)
library(knitr) %>% suppressWarnings()
knitr::opts_chunk$set(echo=F)
theme_set(theme_light())
source('./lib/ncsc_refs.R')

username <- NULL
password <- NULL


db.21 <- referee_database.2021()
db.22 <- referee_database.2022()
referees <- ungroup(filter(db.22, !is.na(Referee)))

```
## Breakdown of Officials

We have a substantial number of referees under the age of 20 with solid representation of women, however the data suggests a substantial attrition with women disproportionately dropping out.

```{r younger, fig.width=6, fig.height=2}

g1 <- ggplot(referees) +
    aes(x=Age, fill=Gender) +
    geom_histogram(breaks = 2.5 + 5*(2:15), color='black', size=0.25) +
    scale_x_continuous(breaks = 5*(2:15)) +
    ylab("Number of Referees") +
    ggtitle("Number of Registered Referees by Age",
            subtitle="Grouped by five years")

g2 <- filter(referees, Age <= 25) |>
    ggplot() +
    aes(x=Age, fill=Gender) +
    geom_histogram(binwidth=1, color='black', size=0.25, show.legend = F) +
    ylab("Number of Referees") +
    scale_x_continuous(breaks=13:25) +
    ggtitle("Young Referees Year to Year") +
    ylab(NULL)

grid.arrange(g1, g2, ncol=2)

```

```{r mapinfo} 
map_theme <- theme_bw() +
    theme(axis.text=element_blank(),
          axis.ticks=element_blank(),
          line=element_blank(),
          plot.background=element_rect(fill='#EEEEEE', size=1, color='#666666'),
          legend.background=element_rect(fill='#E9E9E9',size=0))


portland_metro <- c('Portland', 'Wilsonville', 'Forest Grove', 'Beaverton', 'Oregon City')
statewide <- c('Portland', 'Medford', 'Roseburg', 'Bend', 'Salem', 'Eugene')
cities.or <- read.csv('uscities.csv') |>
    filter(state_id=='OR') |>
    filter(city %in% c(portland_metro, statewide))

cities.statewide <- filter(cities.or, city %in% statewide)
cities.metro <- filter(cities.or, city %in% portland_metro)

```

## Locations of Registered Officials

```{r reflocs, fig.height=5, fig.width=9}
counts <- group_by(referees, Zip) |> summarize(count = n())

# Use tigris lib to get Zip info in sf form
options(tigris_use_cache = TRUE)

# Get zipcodes and join with referee counts by Zip.
Zipcodes.sf <- suppressMessages(sf::st_as_sf(tigris::zctas(year=2010, cb = F, state='OR')))
RefsByZipcodes.sf <- merge(Zipcodes.sf, counts, by.y='Zip', by.x='ZCTA5CE10')

# Get outline of Oregon
states <- suppressMessages(tigris::states(cb=T, class='sf'))
Oregon.sf <- sf::st_geometry(filter(states, NAME=='Oregon'))

rev_sqrt_trans <- function() {
    scales::trans_new(
        name = "rev_sqrt", 
        transform = function(x) -sqrt(abs(x)), 
        inverse = function(x) x^2,
        domain = c(0, Inf))
}

refs.statewide <- ggplot(cities.statewide) +
    geom_sf(data=Oregon.sf) + 
    geom_sf(data=RefsByZipcodes.sf, 
            aes(fill = count), alpha=0.8) +
    geom_point(aes(x=lng, y=lat),
               color='green',
               size=2,
               alpha=0.7,
               show.legend = F) +
    geom_label(aes(x=lng, y=lat, label=city, size=population), 
               show.legend = F,
              nudge_x=0.2, 
              alpha=0.5,
              color='black', 
              fontface='bold',
              hjust='left',
              label.r=unit(0.4, 'lines'),
              label.size=0) +
    xlab(NULL) + ylab(NULL) +
    scale_size_continuous(name='size', range=c(4,7)) +
    scale_fill_viridis_c(option = "plasma", trans =  "rev_sqrt", name="Count")+
    ggtitle("Oregon Referees", subtitle = "Count by zip code") +
    map_theme

refs.portland <- ggplot(cities.metro) +
    geom_sf(data=Oregon.sf) + 
    geom_sf(data=RefsByZipcodes.sf, aes(fill = count), 
            alpha=0.8) +
    # geom_point(aes(x=lng, y=lat),
    #            color='green',
    #            size=2,
    #            alpha=0.7,
    #            show.legend = F) +
    geom_label(aes(x=lng, y=lat, label=city, size=population), 
               show.legend = F,
               nudge_x=0.02, 
               alpha=0.5,
               color='black', 
               fontface='bold',
              label.r=unit(0.4, 'lines'),
              label.size=0) +
    xlab(NULL) + ylab(NULL) +
    scale_size_continuous(name='size', range=c(4,9)) +
    scale_y_continuous(limits=c(45.15, 45.80)) + scale_x_continuous(limits=c(-123.3, -122.05)) +
    scale_fill_viridis_c(option = "plasma", trans =  "rev_sqrt", name="Count")+
    ggtitle("Portland Area Referees",  subtitle = "Count by zip code") +
    map_theme

grid.arrange(refs.statewide, refs.portland, 
             ncol=2)
```

### Mentors and Instructors

```{r mentorsmap, fig.height=5, fig.width=9}
mentors <- filter(db.22, !is.na(Mentor) | !is.na(Assessor) | !is.na(Instructor))
counts <- group_by(mentors, Zip) |> summarize(count = as.integer(n()))

MentorsByZipcodes.sf <- merge(Zipcodes.sf, counts, by.y='Zip', by.x='ZCTA5CE10')

fill_colors <- c(`1`='#FFCCCC', `2`='#DD6666', `3`='#CC4444', `4`='#AA1111')[1:max(counts$count)]

mentors.statewide <- ggplot(cities.statewide) +
    geom_sf(data=Oregon.sf) + 
    geom_sf(data=MentorsByZipcodes.sf, 
            aes(fill = factor(count)), alpha=0.8) +
     geom_label(aes(x=lng, y=lat, label=city, size=population), 
               show.legend = F,
              nudge_x=0.2, 
              alpha=0.5,
              color='black', 
              fontface='bold',
              hjust='left',
              label.r=unit(0.4, 'lines'),
              label.size=0) +
    xlab(NULL) + ylab(NULL) +
    scale_size_continuous(name='size', range=c(4,7)) +
    scale_fill_manual(values=fill_colors, name="Count") +
    ggtitle("Statewide Mentors, Assessors and Instructors", 
            subtitle = "Count of by zip code") +
    map_theme

mentors.metro <- ggplot(cities.metro) +
    geom_sf(data=Oregon.sf) + 
    geom_sf(data=MentorsByZipcodes.sf, aes(fill = factor(count)), 
            alpha=0.8) +
    geom_label(aes(x=lng, y=lat, label=city, size=population), 
               show.legend = F,
              nudge_x=0.02, 
              alpha=0.5,
              color='black', 
              fontface='bold',
              label.r=unit(0.4, 'lines'),
              label.size=0) +
     xlab(NULL) + ylab(NULL) +
    scale_size_continuous(name='size', range=c(4,9)) +
    scale_y_continuous(limits=c(45.15, 45.80)) + scale_x_continuous(limits=c(-123.3, -122.05)) +
    scale_fill_manual(values=fill_colors, name="Count") +
    ggtitle("Portland Area Mentors, Assessors and Instructors", 
            subtitle = "Count by zip code") +
    map_theme

grid.arrange(mentors.statewide, mentors.metro, 
             ncol=2)
```

# Oldest and Youngest

Here are the oldest and youngest referees.


<div id="toplists">
```{r top.oldest}

arrange(referees, desc(Age)) |>
    head(20) |>
    mutate(DOB = strftime(DOB, '%b %Y')) |>
    select(`Full Name`=FullName, Born=DOB, Age, Referee) |>
    kbl( # 'html', 
        escape=FALSE, 
        caption="Twenty Oldest Referees") |>
#    column_spec(1, width = "12em") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = F) 

```


```{r top.youngest}

arrange(referees, Age) |>
    head(20) |>
    mutate(DOB = strftime(DOB, '%b %Y')) |>
    select(`Full Name`=FullName, Born=DOB, Age) |>
    kbl(escape=FALSE, caption = "Twenty Youngest Referees") |>
#    column_spec(1, width = "12em") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = F) 

```
</div>


## Year Over Year Analysis

I don't know if this is valid but I compared the registered referees from mid-2021 to the current list.

```{r historical}
count.current <- nrow(db.22) 
count.previous <- nrow(db.21) 
count.renewed <- length(intersect(db.21$ID, db.22$ID))
percent.renewed <- count.renewed/count.current
count.new <- length(setdiff(db.22$ID, db.21$ID))
count.unrenewed <- length(setdiff(db.21$ID, db.22$ID))

```

In 2021, a total of `r comma(count.previous)` officials registered.

In 2022, so far `r comma(count.current)` officials have registered.

`r comma(count.renewed)` or `r percent(percent.renewed)` have at least partially renewed their registrations.

`r comma(count.unrenewed)` officials have yet to update their registration in 2022.

This year, there are *`r comma(count.new)`* officials who are either new or did not register in 2021.

```{r historical_breakdown}

db.21$Renewed <- db.21$ID %in% db.22$ID

refstatus <- db.21 |>
    mutate(Renewed = ID %in% db.22$ID,
           AgeGroup = ifelse(Age < 16, 'U15',
                             ifelse(Age < 19, 'U19', 'Adult'))) |>
    group_by(Gender, AgeGroup) |>
    summarize(Total=n(), Renewed=sum(Renewed)) |>
    pivot_longer(all_of(c('Total', 'Renewed')),
                 names_to='Registered',
                 values_to='Count') |>
    mutate(AgeGroup = factor(AgeGroup, levels=c('U15', 'U19', 'Adult')),
           Registered = factor(Registered, levels=c('Total', 'Renewed')))
unspecified_gender <- filter(refstatus, Gender == 'U') |> pull(Count) |> sum()
```

In analyzing the attrition by gender, it turns out a substantial number of referees are registered without a gender 
(`r unspecified_gender`).  This seems to be some error in data collection since this phenomenon is not reflected in 2022 data.

So this graph is filtered for referees registered with a gender.

```{r attrition_bygender}

filter(refstatus, Gender != 'U') |>            
ggplot() +
    aes(y=Count, x=Gender, fill=Registered) +
    geom_col(position='dodge') +
    scale_fill_discrete(labels=c("Registered 2021", "Renewed")) +
    labs(fill=NULL, title="Registrations by Gender") +
    ylab('Registered Count')
```

Dividing registrations into age groups you can see the biggest drop off in re-registering seems to be referees between 15 and 18 years old.

```{r attrition_byage}
ggplot(refstatus) +
    aes(y=Count, x=AgeGroup, fill=Registered) +
    geom_col(position='dodge') +
    scale_fill_discrete(labels=c("Registered 2021", "Renewed")) +
    labs(fill=NULL, title="Registrations by Age Group") +
    ylab('Registered Count') 
```
